### 2015-04-15 paper.tex
Fixed references
Tidied everything else up

\begin{abstract}
This report compares and contrasts various copyright licenses which meet the criteria for either free or open source software. Two main categories of licenses are explored: permissive licenses which focus on developer freedom and copyleft licenses which focus on end user freedom. This report concludes that despite the plethora of licenses available, software engineers should familiarize themselves mainly with the MIT license, the Apache license and the GNU Genera Public License and refer other licenses to legal professionals.
\end{abstract}
\newpage
\thispagestyle{empty}
\pagenumbering{arabic}
\setcounter{page}{1}

@@

In the 1990s, due to persistent confusion around the phrase ``free software'', the Open Source Initiative was started, mainly as a way to sell the pragmatic benefits of free software to businesses without the ethical issues attached to nonfree software by the Free Software Foundation.\cite{opensourcehistory} In order for software to be open source, its distribution terms must meet the open source definition:

\subsubsection*{Open Source Definition\cite{opensourcedefinition}}
\begin{enumerate}
\item Free Redistribution (must not be restricted)
\item Source Code (must be included)
\item Derived Works (must be allowed)
\item Integrity of the Author's Source Code (may be maintained if patches may be distributed along side it)
\item No Discrimination Against Persons or Groups
\item No Discrimination Against Fields of Endeavor
\item Distribution of License
\item License Must Not Be Specific to a Product
\item License Must Not Restrict Other Software
\item License Must Be Technology-Neutral
\end{enumerate}

@@

Permissive licenses are the licenses which offer the greatest freedom to developers. Developers are generally free to modify and redistribute the osftware subject to few conditions. The most common condition is including the copyright notice and a liability disclaimer with copies of the software.\cite{permissive}

@@

Apache License, Version 2.0 brought more substantial changes. It removed explicit references to the Apache Software Foundation to make use of the license by others easier and clarified some language so that others could more easily contribute to projects licensed under the Apache License. This included a patent grant from any contributor whose contribution infinged their own patent.\cite{apache2}

@@

The Attribution Assurance License is substantially similar to the BSD License, except that the license is GPG-signed and requires that redistribution include the GPG-signed form of the license. This assures downstream recipients of the software that the terms of the license have not been changed by redistributors.\cite{attrassur}

The Academic Free License includes a copyright and a patent license. It also includes a choice of juridiction clause in favour of the author and includes a provision for recovering legal fees incurred enforcing the terms of the license. It has been deemed redundant due to the creation of the Apache License Version 2.0.\cite{academic}

@@

Copyleft licenses are designed to keep software both free and open. As opposed to permissive licenses, copyleft licenses are modelled around the free software philosophy. A copyleft license guarantees that any derived work must be distributed under the same set of freedoms of the original work \cite{copyleft}. While this licensing restriction can greatly improve collaboration in a development ecosystem, it often leads to incompatibilies with other licenses.

@@

Some of the main features of the GNU GPLv2 include:
\begin{enumerate}
\item The ability to copy and redistribute unmodified source code
\item The ability to modify and redistribute modified source code
\item The ability to distribute a compiled version of the software considering that all copies contain a copyright notice and exclusion of warranty.
\item All modified copies are distributed under the GPLv2 license
\item All compiled programs must be accompanied with the corresponding source code or be provided access to the source code.
\end{enumerate}

\cite{gpl2}

@@

The philosophy behind this change is that you shouldn't be criminalized for writing software, no matter the intent. While this does not forbid DRM in GPLed software, section 3 of the GPLv3 license states that ``the system will not count as an effective technological ``protection'' measure'' \cite{gpl3}. This effectively guarantees that if the author distributes software that breaks the DRM, they will not be threatened laws such as the DMCA.

@@

Finally, GPLv3 aims to avoid discriminatory patent deals. In the past companies like Microsoft and Novel have tried to use their thousands of patents to claim royalties on users for patent infringement. GPLv3 avoid this by providing explicit patent protection from the program's contributors and redistributors \cite{gpl3}.

@@

The Mozilla Public License is a weak copyleft license that aims to strike a balance between the BSD and GPL licenses. By striking this balance, the MPL motives collaboration between business and the open source community. The MPL is used in various open source projects including; Thunderbird, Firefox, and Libreoffice.

@@

The Common Public License is a license produced by IBM which encourages open source development in commercial settings, while still maintaining the ability to use software licensed under separate terms. Unlike the MPL, the CPL is a strong copyleft license, sharing much of the same terminology in the GPL, however it lacks compatibility with all versions of GPL due to ``choice of law'' section which restricts legal disputes to specific courts.

The Eclipse Public License is a slightly modified version of the Common Public License which drops the terms related to patent litigations. It is used in the Eclipse IDE and various other Java related projects including the JVM and Clojure.

The Common Development and Distribution License, produced by Sun Microsystems, is based on the MPL.

The Apple Public Source License was the original license under which Apple's Darwin operating system was developed. While APSL is fully copyleft and the FSF supports the license, they do no longer recommend it. This license was essential in the first stages of Apple, however today most of the software released by the company now use the Apache license.

@@

As a result of this large divide, for free software the decision almost always comes down to the GPLv3 or GPLv2. For open source software, it is ideal to choose a GPL license, but if you cannot ahere to the strict requirements of a copyleft license, it usually comes down to either the MIT, Apache or ``modified'' BSD license.

@@

\bibitem{apache1} The Apache Software Foundation. Apache License, Version 1.0 \url{https://www.apache.org/licenses/LICENSE-1.0}

\bibitem{apache11} The Apache Software Foundation. Apache License, Version 1.1 \url{https://www.apache.org/licenses/LICENSE-1.1}

\bibitem{apache2} The Apache Software Foundation. Apache License, Version 2.0 \url{https://www.apache.org/licenses/LICENSE-2.0}

\bibitem{boost} Boost C++ Libraries. Boost Software License. \url{http://www.boost.org/users/license.html}

\bibitem{nolicense} ChooseALicense.com. No License. \url{http://choosealicense.com/no-license/}

\bibitem{billyg} DigiBarn Computer Museum. Bill Gates' Open Letter To Hobbyists. \url{http://www.digibarn.com/collections/newsletters/homebrew/V2_01/gatesletter.html}

\bibitem{originalbsd} Free Software Directory. License:BSD 4Clause. \url{directory.fsf.org/wiki/License:BSD_4Clause}

\bibitem{simplifiedbsd} Free Software Directory. License:FreeBSD. \url{http://directory.fsf.org/wiki?title=License:FreeBSD}

\bibitem{x11} Free Software Directory. License:X11. \url{http://directory.fsf.org/wiki/License:X11}

\bibitem{fsfhistory} GNU Operating System. About the GNU Project \url{https://www.gnu.org/gnu/thegnuproject.html}

\bibitem{netbsd} GNU Operating System. BSD License Problem. \url{https://www.gnu.org/philosophy/bsd.html}

\bibitem{gpl2} GNU Operating System. GNU General Public License, version 2. \url{https://www.gnu.org/licenses/gpl-2.0.html}.

\bibitem{licenselist} GNU Operating System. Various Licenses and Comments About Them. \url{https://www.gnu.org/licenses/license-list.html}

\bibitem{copyleft} GNU Operating System. What is Copyleft? \url{https://www.gnu.org/copyleft/}.

\bibitem{fourfreedoms} GNU Operating System. What is Free Software? \url{https://www.gnu.org/philosophy/free-sw.html}

bibitem{gpl3} GNU Operating System. A Quick Guide to GPLv3. \url{https://www.gnu.org/licenses/quick-guide-gplv3.html}.

\bibitem{openbsd} OpenBSD. OpenBSD Copyright Policy. \url{http://www.openbsd.org/policy.html}

\bibitem{academic} Open Source Initiative. Academic Free License 3.0. \url{http://opensource.org/licenses/AFL-3.0}

\bibitem{attrassur} Open Source Initiative. Attribution Assurance License. \url{http://opensource.org/licenses/AAL}

\bibitem{education} Open Source Initiative. Education Community License, Version 2.0 \url{http://opensource.org/licenses/ECL-2.0}

\bibitem{fair} Open Source Initiative. Fair License (Fair). \url{http://opensource.org/licenses/Fair}

\bibitem{permissive} Open Source Initiative. Frequently Answered Questions. \url{http://opensource.org/faq}

\bibitem{opensourcehistory} Open Source Initiative. History of the OSI. \url{http://opensource.org/history}

\bibitem{mit} Open Source Initiative. The MIT License (MIT). \url{http://opensource.org/licenses/MIT}

\bibitem{opensourcedefinition} Open Source Initiative. The Open Source Definition. \url{http://opensource.org/definition}

\bibitem{sqlite} SQLite. SQLite Copyright. \url{https://www.sqlite.org/copyright.html}

\bibitem{copyright} Teaching Copyright. Copyright Frequently Asked Questions. \url{http://www.teachingcopyright.org/handout/copyright-faq}

\bibitem{wtfpl} WTFPL – Do What the Fuck You Want to Public License. Frequently Asked Questions. \url{http://www.wtfpl.net/faq/}

\bibitem{lgpl} GNU Operating System. GNU Lesser General Public License. \url{https://www.gnu.org/licenses/lgpl.html}.

\bibitem{agpl} GNU Operating System. Why the Affero GPL. \url{https://gnu.org/licenses/why-affero-gpl.html}.

\bibitem{mpl} Mozilla. MPL 2.0 FAQ. \url{https://www.mozilla.org/MPL/2.0/FAQ.html}

\bibitem{license-stats} Black Duck. Top 20 Open Source Licenses. \url{https://www.blackducksoftware.com/resources/data/top-20-open-source-licenses}.

>>>>

\begin{abstract}
This report compares and contrasts various copyright licenses which meet the criteria for either free or open source software. Two main categories of licenses are explored: permissive licenses which focus on developer freedom and copyleft licenses which focus on end user freedom. This report concludes that despite the plethora of licenses available, software engineers should familiarize themselves mainly with the MIT license, the Apache license and the GNU General Public License and refer other licenses to legal professionals.
\end{abstract}
\newpage
\thispagestyle{empty}
\pagenumbering{arabic}
\setcounter{page}{1}

@@

In the 1990s, due to persistent confusion around the phrase ``free software'', the Open Source Initiative was started, mainly as a way to sell the pragmatic benefits of free software to businesses without the ethical issues attached to non-free software by the Free Software Foundation.\cite{opensourcehistory} In order for software to be open source, its distribution terms must meet the open source definition:

\subsubsection*{Open Source Definition\cite{opensourcedefinition}}
\begin{enumerate}
\item Free Redistribution (must not be restricted)
\item Source Code (must be included)
\item Derived Works (must be allowed)
\item Integrity of the Author's Source Code (may be maintained if patches may be distributed along side it)
\item No Discrimination Against Persons or Groups
\item No Discrimination Against Fields of Endeavour
\item Distribution of License
\item License Must Not Be Specific to a Product
\item License Must Not Restrict Other Software
\item License Must Be Technology-Neutral
\end{enumerate}

@@

Permissive licenses are the licenses which offer the greatest freedom to developers. Developers are generally free to modify and redistribute the software subject to few conditions. The most common condition is including the copyright notice and a liability disclaimer with copies of the software.\cite{permissive}

@@

Apache License, Version 2.0 brought more substantial changes. It removed explicit references to the Apache Software Foundation to make use of the license by others easier and clarified some language so that others could more easily contribute to projects licensed under the Apache License. This included a patent grant from any contributor whose contribution infringed their own patent.\cite{apache2}

@@

The Attribution Assurance License is substantially similar to the BSD License, except that the license is GPG-signed and requires that redistribution include the GPG-signed form of the license. This assures downstream recipients of the software that the terms of the license have not been changed by redistributor.\cite{attrassur}

The Academic Free License includes a copyright and a patent license. It also includes a choice of jurisdiction clause in favour of the author and includes a provision for recovering legal fees incurred enforcing the terms of the license. It has been deemed redundant due to the creation of the Apache License Version 2.0.\cite{academic}

@@

Copyleft licenses are designed to keep software both free and open. As opposed to permissive licenses, copyleft licenses are modelled around the free software philosophy. A copyleft license guarantees that any derived work must be distributed under the same set of freedoms of the original work \cite{copyleft}. While this licensing restriction can greatly improve collaboration in a development ecosystem, it often leads to incompatibilities with other licenses.

@@

Some of the main features of the GNU GPLv2 include \cite{gpl2}:
\begin{enumerate}
\item The ability to copy and redistribute unmodified source code
\item The ability to modify and redistribute modified source code
\item The ability to distribute a compiled version of the software considering that all copies contain a copyright notice and exclusion of warranty.
\item All modified copies are distributed under the GPLv2 license
\item All compiled programs must be accompanied with the corresponding source code or be provided access to the source code.
\end{enumerate}

@@

The philosophy behind this change is that you shouldn't be criminalized for writing software, no matter the intent. While this does not forbid DRM in GPLed software, section 3 of the GPLv3 license states that ``the system will not count as an effective technological ``protection'' measure'' \cite{gpl3}. This effectively guarantees that if the author distributes software that breaks the DRM, they will not be threatened by laws such as the DMCA.

@@

Finally, GPLv3 aims to avoid discriminatory patent deals. In the past companies like Microsoft and Novell have tried to use their thousands of patents to claim royalties on users for patent infringement. GPLv3 avoids this by providing explicit patent protection from the program's contributors and redistributors \cite{gpl3}.

@@

The Mozilla Public License is a weak copyleft license that aims to strike a balance between the BSD and GPL licenses. By striking this balance, the MPL motives collaboration between business and the open source community. The MPL is used in various open source projects including; Thunderbird, Firefox, and LibreOffice.

@@

The Common Public License is a license produced by IBM which encourages open source development in commercial settings, while still maintaining the ability to use software licensed under separate terms. Unlike the MPL, the CPL is a strong copyleft license, sharing much of the same terminology in the GPL, however it lacks compatibility with all versions of GPL due to ``choice of law'' section which restricts legal disputes to specific courts \cite{cpl}.

The Eclipse Public License is a slightly modified version of the Common Public License which drops the terms related to patent litigations. It is used in the Eclipse IDE and various other Java related projects including the JVM and Clojure \cite{epl}.

The Apple Public Source License was the original license under which Apple's Darwin operating system was developed. While APSL is fully copyleft and the FSF supports the license, they do no longer recommend it \cite{licenselist}. This license was essential in the first stages of Apple, however today most of the software released by the company now use the Apache license.

@@

As a result of this large divide, for free software, the decision almost always comes down to the GPLv3 or GPLv2. For open source software, it is ideal to choose a GPL license, but if you cannot adhere to the strict requirements of a copyleft license, it usually comes down to either the MIT, Apache or ``modified'' BSD license.

@@

\raggedright
\bibitem{apache1} The Apache Software Foundation. Apache License, Version 1.0 [Online]. Available: \url{https://www.apache.org/licenses/LICENSE-1.0}

\bibitem{apache11} The Apache Software Foundation. Apache License, Version 1.1 [Online]. Available: \url{https://www.apache.org/licenses/LICENSE-1.1}

\bibitem{apache2} The Apache Software Foundation. Apache License, Version 2.0 [Online]. Available: \url{https://www.apache.org/licenses/LICENSE-2.0}

\bibitem{license-stats} Black Duck. Top 20 Open Source Licenses [Online]. Available: \url{https://www.blackducksoftware.com/resources/data/top-20-open-source-licenses}.

\bibitem{boost} Boost C++ Libraries. Boost Software License [Online]. Available: \url{http://www.boost.org/users/license.html}

\bibitem{nolicense} ChooseALicense.com. No License [Online]. Available: \url{http://choosealicense.com/no-license/}

\bibitem{billyg} DigiBarn Computer Museum. Bill Gates' Open Letter To Hobbyists [Online]. Available: \url{http://www.digibarn.com/collections/newsletters/homebrew/V2_01/gatesletter.html}

\bibitem{epl} Eclipse. Eclipse Public License (EPL) Frequently Asked Questions [Online]. Available: \url{http://www.eclipse.org/legal/eplfaq.php#CPLEPL}

\bibitem{originalbsd} Free Software Directory. License:BSD 4Clause [Online]. Available: \url{directory.fsf.org/wiki/License:BSD_4Clause}

\bibitem{simplifiedbsd} Free Software Directory. License:FreeBSD [Online]. Available: \url{http://directory.fsf.org/wiki?title=License:FreeBSD}

\bibitem{x11} Free Software Directory. License:X11 [Online]. Available: \url{http://directory.fsf.org/wiki/License:X11}

\bibitem{fsfhistory} GNU Operating System. About the GNU Projec [Online]. Available: \url{https://www.gnu.org/gnu/thegnuproject.html}

\bibitem{netbsd} GNU Operating System. BSD License Problem [Online]. Available: \url{https://www.gnu.org/philosophy/bsd.html}

\bibitem{gpl2} GNU Operating System. GNU General Public License, version 2 [Online]. Available: \url{https://www.gnu.org/licenses/gpl-2.0.html}.

\bibitem{licenselist} GNU Operating System. Various Licenses and Comments About Them [Online]. Available: \url{https://www.gnu.org/licenses/license-list.html}

\bibitem{copyleft} GNU Operating System. What is Copyleft [Online]? Available: \url{https://www.gnu.org/copyleft/}.

\bibitem{fourfreedoms} GNU Operating System. What is Free Software [Online]? Available: \url{https://www.gnu.org/philosophy/free-sw.html}

\bibitem{gpl3} GNU Operating System. A Quick Guide to GPLv3 [Online]. Available: \url{https://www.gnu.org/licenses/quick-guide-gplv3.html}.

\bibitem{lgpl} GNU Operating System. GNU Lesser General Public License [Online]. Available: \url{https://www.gnu.org/licenses/lgpl.html}.

\bibitem{agpl} GNU Operating System. Why the Affero GPL [Online]. Available: \url{https://gnu.org/licenses/why-affero-gpl.html}.

\bibitem{mpl} Mozilla. MPL 2.0 FAQ [Online]. Available: \url{https://www.mozilla.org/MPL/2.0/FAQ.html}

\bibitem{openbsd} OpenBSD. OpenBSD Copyright Policy [Online]. Available: \url{http://www.openbsd.org/policy.html}

\bibitem{academic} Open Source Initiative. Academic Free License 3.0 [Online]. Available: \url{http://opensource.org/licenses/AFL-3.0}

\bibitem{attrassur} Open Source Initiative. Attribution Assurance License [Online]. Available: \url{http://opensource.org/licenses/AAL}

\bibitem{cpl} Open Source Initiative. Common Public License Version 1.0 [Online]. Available: \url{http://opensource.org/licenses/CPL-1.0}

\bibitem{education} Open Source Initiative. Education Community License, Version 2.0 [Online]. Available: \url{http://opensource.org/licenses/ECL-2.0}

\bibitem{fair} Open Source Initiative. Fair License (Fair) [Online]. Available: \url{http://opensource.org/licenses/Fair}

\bibitem{permissive} Open Source Initiative. Frequently Answered Questions [Online]. Available: \url{http://opensource.org/faq}

\bibitem{opensourcehistory} Open Source Initiative. History of the OSI [Online]. Available: \url{http://opensource.org/history}

\bibitem{mit} Open Source Initiative. The MIT License (MIT) [Online]. Available: \url{http://opensource.org/licenses/MIT}

\bibitem{opensourcedefinition} Open Source Initiative. The Open Source Definition [Online]. Available: \url{http://opensource.org/definition}

\bibitem{sqlite} SQLite. SQLite Copyright [Online]. Available: \url{https://www.sqlite.org/copyright.html}

\bibitem{copyright} Teaching Copyright. Copyright Frequently Asked Questions [Online]. Available: \url{http://www.teachingcopyright.org/handout/copyright-faq}

\bibitem{wtfpl} WTPF. WTFPL – Do What the Fuck You Want to Public License. Frequently Asked Questions [Online]. Available: \url{http://www.wtfpl.net/faq/}
